|

### 2021-09-24 paper.Rmd
mostly formatting

The experimental treatment consists of a combination of two types of information, each mimicking a short online news article. One type of information reports the results of a study claiming that genetically modified corn lead to increased cancer rates in rats, and features the headline *"GM corn-fed rats die earlier"*. The wording is partly taken from actual news coverage of the study in 2012, and partly from treatment material in @clifford2019emotional. The respective study has been retracted due to non-adherence to scientific procedures, and the current scientific evidence does not point to severe health problems associated with genetically modified animal feed [@brody2018are]. The information within the study can thus be classified as misinformation.^[This is consistent with a definition of resulting mis*perceptions* as "factual beliefs that are false or contradict the best available evidence in the public domain" [@flynn2017nature, p.128]. If one would presume the intent to mislead, this would then be classified as disinformation.] I vary this misinformation to be either low or high in emotional framing by ex- or including graphical images of a rat with a cancer tumor and rotten corn, images that are taken from @clifford2019emotional and shown to induce negative feelings, specifically disgust.

@@

* In the experiment, I pictures as visual cues to convey emotional frames. While this manipulation seems to be working for the intended emotion of disgust, I cannot distinguish between the effect of emotional framing through images and the effect of visual cues per se. Research on the effects of visuals for dis- and corrective information shows that visual disinformation is only slightly more credible whereas visual correction is not more effective than textual correction [@hameleers2020picture]. As an explanation, the authors specualte that the textual format works because it "fits fact-checkers' reliance on short, rational and eveidecne-based argumentation" [@hameleers2020picture, p. 297]. In their study, imapges are conceptualized as providing "indexality", i.e. a clearer link to reality, and not emotionality as in my study. Overall, the non-existence of an effect of visual cues as "indexality" on the effectiveness of correction provides support for my argument that in my study, it is the emotionality of the images and not the visual cues per se that make for the different effects, but further research is needed to completely disentangle these effects.

>>>>

The experimental treatment consists of a combination of two types of information, each mimicking a short online news article. One type of information reports the results of a study claiming that genetically modified corn lead to increased cancer rates in rats, and features the headline *"GM corn-fed rats die earlier"*. The wording is partly taken from actual news coverage of the study in 2012, and partly from treatment material in @clifford2019emotional. The respective study has been retracted due to non-adherence to scientific procedures, and the current scientific evidence does not point to severe health problems associated with genetically modified animal feed [@brody2018are]. The information within the study can thus be classified as misinformation.^[This is consistent with a definition of resulting mis*perceptions* as "factual beliefs that are false or contradict the best available evidence in the public domain" [@flynn2017nature, p.128]. It should be noted, however, that the treatment material in @clifford2019emotional uses this example only as one of several and remains vague without giving specific scientific information.] I vary this misinformation to be either low or high in emotional framing by ex- or including graphical images of a rat with a cancer tumor and rotten corn, images that are taken from @clifford2019emotional and shown to induce negative feelings, specifically disgust.

@@

* In the experiment, I use pictures as visual cues to convey emotional frames. While this manipulation seems to be working for the intended emotion of disgust, I cannot distinguish between the effect of emotional framing through images and the effect of visual cues per se. Research on the effects of visuals for dis- and corrective information shows that visual disinformation is only slightly more credible whereas visual correction is not more effective than textual correction [@hameleers2020picture]. As an explanation, the authors specualte that the textual format works because it "fits fact-checkers' reliance on short, rational and evidence-based argumentation" [@hameleers2020picture, p. 297]. In their study, imapges are conceptualized as providing "indexicality", i.e. a clearer link to reality, and not emotionality as in my study. Overall, the non-existence of an effect of visual cues as "indexicality" on the effectiveness of correction provides support for my argument that in my study, it is the emotionality of the images and not the visual cues per se that make for the different effects, but further research is needed to completely disentangle these effects.
